---
title: "VisPaper_Analaysis"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
```

# Data Preparation
```{r cars}
#keywords
viskeywords<-readxl::read_xlsx("../Tables/datascie_vis_alignement.xlsx")
processOrder<-unique(viskeywords$Process)

#vispubdata 
vispubdata_to18<-readxl::read_xlsx("data/IEEE VIS papers 1990-2018.xlsx",sheet = 1)
vispubdata_19<-readxl::read_xlsx("data/IEEE VIS papers 2019.xlsx")

vispubdata<-rbind(vispubdata_to18,vispubdata_19) %>%
  dplyr::filter(!is.na(Abstract) & !is.na(AuthorKeywords)) %>%
  mutate(Abstract = tolower(Abstract)) %>%
  mutate(AuthorKeywords = tolower(AuthorKeywords))
```

```{r, echo=FALSE}

viskeywords<-viskeywords %>%
  tidyr::separate_rows(keyTerms,sep=";") %>%
  dplyr::mutate(keyTerms = trimws(keyTerms))%>%
  dplyr::filter(keyTerms != "")

vispubkey <- vispubdata %>%
  dplyr::select(DOI,AuthorKeywords) %>%
  tidyr::separate_rows(AuthorKeywords,sep=",") %>%
  dplyr::mutate(AuthorKeywords = tolower(trimws(AuthorKeywords)))

keywordsJoint<-full_join(viskeywords,vispubkey,by = c("keyTerms" = "AuthorKeywords"))
  
```

# Examining Key Terms

Looking athte terms from keyvis.org dataset that was precompiled. Emphasizing the top ten words.
```{r}
keywordsJoint %>%
  dplyr::filter(keyTerms %in% viskeywords$keyTerms)%>%
  group_by(keyTerms) %>%
  tally() %>%
  arrange(-n)

uniquePapers<-keywordsJoint %>%
  dplyr::filter(keyTerms %in% viskeywords$keyTerms) %>%
  select(DOI)%>%
  filter(!is.na(DOI))%>%
  unique() %>%
  count()

uniquePapers<-uniquePapers$n
totalPapers<-nrow(vispubdata)
totalPer = round((uniquePapers/totalPapers) * 100,2)
```

There are `r uniquePapers` from a total `r totalPapers` (`r totalPer`%) that match these papers. Most of these (185) (`r round((185/totalPapers)*100,2)`%) are visual analytics, most likely due to the VAST conference.


Now, look at the higher and lower order processes involed.

```{r}
process_coverag<-keywordsJoint %>%
  dplyr::filter(keyTerms %in% viskeywords$keyTerms) %>%
  right_join(viskeywords) %>%
  left_join(vispubdata)

process_coverag$higherOrderProcess <- factor(as.character(process_coverag$higherOrderProcess),
                                           levels = c("Preparation","Analysis","Communication"))

process_coverag$Process<-factor(as.character(process_coverag$Process), 
                                 levels= processOrder)

ggplot(data=process_coverag,aes(x=Process))+
  geom_bar()+
  coord_flip()+
  theme_bw()
```

Now... put this on my process chart

```{r}
rects <- data.frame(x = c(seq(1,24,by=6),seq(1,24,by=6),seq(1,12,by=6)),
                    y = c(rep(6,4),rep(3,4),rep(0,2)),
                    text = processOrder)

process_coverag %>%
  group_by(Process) %>%
  tally()%>%
  mutate(n = n/nrow(process_coverag))%>%
  full_join(rects,by=c("Process" = "text")) %>%
  ggplot(aes(x, y, fill = n, label = Process)) +
    geom_tile(width = 4.5, height = 2,color = "black") + # make square tiles
    geom_text(color = "black") + # add white text in the middle
    scale_fill_gradient(high = "black",low="lightgrey",na.value = "white", limits = c(0,1),label=percent)+
    #scale_fill_identity(guide = "none") + # color the tiles with the colors in the data frame
    coord_fixed() + # make sure tiles are square
    theme_void() # remove any axis markings
```

# Papers

## Analysis

### Modeling
```{r}
modeling<-process_coverag %>%
  filter(Process == "Modeling") %>%
  mutate(citation = as.numeric(`XploreCitationCount-2020-01`))
```

## Communication 
### Collaboration
```{r}
collab<-process_coverag %>%
  filter(Process == "Collaboration") %>%
  mutate(citation = as.numeric(`XploreCitationCount-2020-01`))
```